---
title: "Lab4: Avocado Prices"
author: "Sebastian Boksem"
format: html
editor: visual
---

## Declaring Libraries

```{r}
#| execute:
#| output: FALSE
library(tidyverse)
```

## Importing Data Set

```{r}
#| execute:
#| output: FALSE
avocado <- read_csv(here::here("supporting_artifacts", "Lab4", "avocado.csv"))

```

```{r}
dim(avocado)
str(avocado)
```

This data set seems to contain the total volume of three different price lookup's of avocados sold in different stores. In addition to this, the data, type (whether it is organic or conventional), the year, region, and average price is included. There seems to be character, dates, and non-integer data type values. Overall, there seems to be 18249 rows with 14 columns. 4046 represents Small Hass, 4225 represents Large Hass, 4770 represents extra large avocados.

```{r}
major_region <- c('Northeast', 'Southeast', 'SouthCentral', 'Plains', 'West', 'Midsouth', 'GreatLakes')
major_state <- c('NewYork', 'California', 'SouthCarolina', 'WestTexNewMexico')
california_cities <- c('LosAngeles', 'SanDiego', 'Sacramento', 'SanFrancisco')
total_us <- c('TotalUS')
major_region <- data.frame(major_region)
major_state <- data.frame(major_state)
california_cities <- data.frame(california_cities)
total_us <- data.frame(total_us)
avocado_major_region <- avocado |> 
  semi_join(major_region, by = c('region' = 'major_region'))
avocado_no_major_region <- avocado |> 
  anti_join(major_region, by = c('region' = 'major_region'))
avocado_no_major_region_or_state <- avocado_no_major_region |>
  anti_join(major_state, by = c('region' = 'major_state'))
avocado_metro_cities <- avocado_no_major_region_or_state |>
  anti_join(total_us, by = c('region' = 'total_us'))
avocado_total_us <- avocado |>
  semi_join(total_us, by = c('region' = 'total_us'))
avocado_california <- avocado |>
  semi_join(california_cities, by = c('region' = 'california_cities'))


```

**REFELCTION:**

**Upon looking back on this question, I think the overall application of chaining and piping all the anti joins together is important because it shows how exactly this joins can be used to clean data in a faster way. Also it is important to include the states Minnesota and Michigan and I've added the great lakes to the region. For the future I will chain the anti-joins to optimize the code. I've added a new section below in which I have done this:**

```{r}
avocado |> 
  anti_join(major_region, by = c('region' = 'major_region')) |>
  anti_join(major_state, by = c('region' = 'major_state')) |>
  filter(region != 'TotalUS') |>
  rename("Small" =  `4046`,
         "Large" =  `4225`,
         "Extra Large" = `4770`)
```

## 3.

```{r}
avocado_major_region |> 
  filter(
    type == 'organic',
    year == 2017
  ) |>
  group_by(region) |>
  summarize(total = sum(`4046`)) |>
slice_max(total)
  
```

It appears that the West sold the most organic, small Hass avocados in 2017. They sold 1,870,206 avocados.

## 4.

```{r}
avocado_total_us |>
  separate(col = Date, into = c('Year', 'Month', 'Day'), sep = '-') |>
  group_by(Month) |>
  summarize(total = sum(as.numeric(`Total Volume`))) |>
slice_max(total)

  
```

It appears that month 2, or February, has sold the most volume of avocados. They sold 665,625,110 avocados.

## 5.

```{r}
avocado_metro_cities |>
  group_by(region) |>
  summarize(average = mean(`Total Volume`)) |>
  slice_max(average, n = 5)  |>
  select(region) |>
  left_join(avocado, by = 'region') |>
  ggplot( 
       mapping = aes(y = reorder(region, `Total Volume`, median),
                     x = `Total Volume`,
                     ) 
                     ) + geom_boxplot() +
  labs(x = "Total Volume", y = "Metro Region")


```

## 6.

```{r}
#avocado_california |> 
# group_by(region, type) |>
#  summarize(avg = mean(AveragePrice))  |>
#  pivot_wider(names_from = type, values_from = avg) |>
#  mutate(difference = conventional - organic)
            
avocado_california |> 
  group_by(region, type) |>
  summarize(avg = mean(AveragePrice)) |>
  select(region, type) |>
  left_join(avocado, by = c('region' = 'region', 'type' = 'type')) |>
   ggplot( 
       mapping = aes(y = AveragePrice,
                     x = region,
                     color = type
                     ) 
                     ) + geom_boxplot()


#avocado_california |> 
#  group_by(region, type) |>
#  summarize(median_avocado = median(AveragePrice))  |>
#  pivot_wider(names_from = type, values_from = median_avocado) |>
#  mutate(difference = conventional - organic)




```

From this graphs showing both the difference from a median and mean standpoint, it seems like San Francisco would have the higher difference in conventional versus organic prices. For the mean , there is a 91 cent difference while for the median there is a 81 cent difference.

## 7.

```{r}
avocado_california |>
  rename("Small" =  `4046`,
         "Large" =  `4225`,
         "Extra Large" = `4770`) |>
  mutate(total = Small + Large + `Extra Large`,
         Small_Prop = Small / total,
         Large_Prop = Large / total,
        EXL_Prop = `Extra Large` / total) |>
  group_by(region,type) |>
  summarize(Small_Prop = mean(Small_Prop),
            Large_Prop = mean(Large_Prop),
            Extra_Large_Prop = mean(EXL_Prop)) |>
pivot_longer(`Small_Prop`:`Extra_Large_Prop`,
               names_to = "Size",
               values_to = "Avocados Prop") |>
  ggplot() + geom_col(mapping = aes(x = region,y = `Avocados Prop`, fill = Size)) + facet_grid(. ~ type) +
  theme(axis.text.x = element_text(angle = 90))
```
