---
title: "Lab4: Avocado Prices"
author: "Sebastian Boksem"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
editor: visual
---

## From Lab4

**REFLECTION:**

**I have removed the unnecessary output from the first code chunks. I feel like this has overall taught me more about how YAML format works within code chunks and initially I was just adding it in the first YAML above. I will now know not to add unnecessary output in future labs.**

```{r}
#| execute:
#| output: FALSE
library(tidyverse)
```

Importing Data Set

```{r}
#| execute:
#| output: FALSE
avocado <- read_csv(here::here("supporting_artifacts", "Lab4", "avocado.csv"))
```

```{r}
major_region <- c('Northeast', 'Southeast', 'SouthCentral', 'Plains', 'West', 'Midsouth')
major_state <- c('NewYork', 'California', 'SouthCarolina', 'WestTexNewMexico')
california_cities <- c('LosAngeles', 'SanDiego', 'Sacramento', 'SanFrancisco')
total_us <- c('TotalUS')
major_region <- data.frame(major_region)
major_state <- data.frame(major_state)
california_cities <- data.frame(california_cities)
avocado_california <- avocado |>
  semi_join(california_cities, by = c('region' = 'california_cities'))

```

## Challenge

```{r}
#Custom Dataset with median house price taken from Google, 2018 housing prices due to data in Avocados ending there. https://www.redfin.com/city/11203/CA/Los-Angeles etc.

City <- c('Sacramento', 'LosAngeles','SanFrancisco','SanDiego')
Price <- c(330000, 705000, 1402000, 600000)
HousePrice <- data.frame(City, Price)
avocado_california <- avocado |>
  semi_join(california_cities, by = c('region' = 'california_cities'))


avocado_california |>
select(AveragePrice, `Total Volume`, type, region) |>
  group_by(region, type) |>
  summarize(med_avocado_price = median(AveragePrice)) |>
   left_join(HousePrice, by = c("region" = "City")) |>
   ggplot( 
       mapping = aes(y = med_avocado_price,
                     x = Price,
                     color = type
                     ) 
                     ) + geom_point() +
  labs(x = "Housing Price", y = "Median Avocado_Price")


```

It seems that there seems to be somewhat a positive correlation between housing prices and avocado prices. However, it is quite difficult to tell as there is not a lot of data supported from the housing prices in every city. This could also be associated with cost of living being higher, and so items like Avocados would naturally be a higher price.
